Peer-to-Peer 6502 Cellular Automata

### 2018-07-03 paper.tex
cube

@@

The blockchain was introduced for cryptocurrencies \cite{Nakamoto2008}.
It can also be used to drive computation on a distributed virtual machine.
In the Ethereum network, this is used to implement smart contracts.
Here we describe a different (simpler) use: the simulation of a virtual world of generic cellular automata.

@@

After some experiments with 32-bit ARM RISC architectures,
I settled on the 8-bit MOS Technology 6502 processor for a CPU.
The corresponding 16-bit memory map is shown in Table~\ref{tab:MemoryMap}.

In this memory map, the 27 pages of storage
for the cell and its 26 neighbors
in the $3 \times 3 \times 3$ unit cube centered on $(x,y,z)$
(in cellular automata terminology, the {\em three-dimensional Moore neighborhood})
are mapped directly to bytes in the address range \hexrange{4000}{7FFF} of the virtual machine.
The detailed description of this part of the memory map
is given in Table~\ref{tab:MooreNeighborhood}.

@@

\begin{table}
\begin{tabular}{lll}
  \hline
  Byte & Bit(s) & Meaning \\
  \hline
  MSB & 4,5 & $i$ \\
  MSB & 2,3 & $j$ \\
  MSB & 0,1 & $k$ \\
  LSB & 0-7 & Address within page \\
  \hline
\end{tabular}
\caption{
  Interpretation of a 16-bit address in the memory-mapped range \hexrange{4000}{7FFF}
  as pointing to the cell $(x+i-1,y+j-1,z+k-1)$.
  See Table~\ref{tab:MooreNeighborhood} and Table~\ref{tab:VonNeumannNeighborhood}.
  Note that $i,j,k, \in \{ 0,1,2 \}$.
  \label{tab:AddressBits}
}
\end{table}

>>>>

\usepackage{graphicx}

@@

The blockchain was introduced as a distributed ledger for cryptocurrencies \cite{Nakamoto2008}.
More generally it can be used to drive computation on a distributed virtual machine.
In the Ethereum network, the virtual machine is used to implement smart contracts.
Here we describe a different (simpler) use: the simulation of a virtual world of generic cellular automata.

@@

After some experiments with 32-bit ARM RISC architectures,
I settled on the 8-bit MOS Technology 6502 processor for a CPU.
The corresponding 16-bit memory map is shown in Table~\ref{tab:MemoryMap} and Figure~\ref{fig:UnitCube}.

In this memory map, the 27 pages of storage
for the cell and its 26 neighbors
in the $3 \times 3 \times 3$ unit cube centered on $(x,y,z)$
(in cellular automata terminology, the {\em three-dimensional Moore neighborhood}, shown in Figure~\ref{fig:UnitCube})
are mapped directly to bytes in the address range \hexrange{4000}{7FFF} of the virtual machine.
The detailed description of this part of the memory map
is given in Table~\ref{tab:MooreNeighborhood}.

@@

\begin{table}
\begin{tabular}{lll}
  \hline
  Byte & Bit(s) & Meaning \\
  \hline
  MSB & 4,5 & $i$ \\
  MSB & 2,3 & $j$ \\
  MSB & 0,1 & $k$ \\
  LSB & 0-7 & Address within page \\
  \hline
\end{tabular}
\caption{
  Interpretation of a 16-bit address in the memory-mapped range \hexrange{4000}{7FFF}
  as pointing to the cell $(x+i-1,y+j-1,z+k-1)$.
  See Table~\ref{tab:MooreNeighborhood} and Table~\ref{tab:VonNeumannNeighborhood}, and Figure~\ref{fig:UnitCube}.
  Note that $i,j,k, \in \{ 0,1,2 \}$.
  \label{tab:AddressBits}
}
\end{table}

\begin{figure}
  \includegraphics[width=\textwidth]{unitcube.pdf}
\caption{
  The unit cube of the Moore neighborhood, and its relationship to the memory map of Table~\ref{tab:MooreNeighborhood}.
  The central cell is at $(1,1,1)$, and is mapped to zero page.
  All other cells are mapped to pages in the range \hexrange{4000}{7FFF},
  the first six bits of the address MSB corresponding to the $x$, $y$ and $z$ co-ordinates within this cell (see Table~\ref{tab:AddressBits}).
  \label{fig:UnitCube}
}
\end{figure}
